---
title: "Regression"
output: github_document
---


```{r}
library(tidyverse)
library(sf)
library(tmap)
library(spgwr)
library(jtools)
library(tmaptools)
library(viridis)
library(modelr)
library(mgcv)
library(ggridges)
library(patchwork)
library(lubridate)
library(splitstackshape)
library(stringr)
library(naniar)
library(dplyr) 
library(broom)
library(AER)


knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 10, 
  fig.height = 10,
  out.width = "90%")

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)


scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
health_outcomes = 
  read_csv("./data/PLACES_County_2022.csv") %>% 
  janitor::clean_names() %>% 
  rename(county = county_name, state = state_abbr) %>% 
  mutate(
    county = as.factor(county)
  ) %>% 
  select(state, county, casthma_crude_prev,casthma_crude95ci, cancer_crude_prev,cancer_crude95ci,access2_crude_prev,depression_crude_prev,csmoking_crude_prev,total_population) %>% 
  filter(state == "PA") %>%
  mutate(asthma_ci = str_remove_all(casthma_crude95ci, '[()]'),
         cancer_ci = str_remove_all(cancer_crude95ci, '[()]')) %>%
  select(-casthma_crude95ci,-cancer_crude95ci) %>%
  separate(asthma_ci, into = c("asthma_lower", "asthma_upper"), sep = ",") %>%
   separate(cancer_ci, into = c("cancer_lower", "cancer_upper"), sep = ",") %>%
  mutate(
    asthma_lower = as.numeric(asthma_lower),
    asthma_upper = as.numeric(asthma_upper),
    cancer_lower = as.numeric(cancer_lower),
    cancer_upper = as.numeric(cancer_upper)
  )

```

```{r}
superfund_sites =
  read_csv("./data/superfund_NPL_sites.csv") %>% 
  janitor::clean_names() %>% 
  select(-region_id,-construction_completion_number, -site_listing_narrative, -site_progress_profile, -ends_with("notice"), -restoration_fr_notice_jumper_page, -noid_date, -deletion_date,-x,-y,-state) %>% 
  filter(status == "NPL Site")

superfund_by_county =
  superfund_sites %>% 
  group_by(county) %>% 
  summarize(n_superfund = n()) %>% 
  mutate(
    county = as.factor(county)) %>%
  as.data.frame(superfund_by_county)
```

```{r}
combined_data_1 =
  left_join(health_outcomes, superfund_sites, by = "county") 

combined_data_final=left_join(combined_data_1, superfund_by_county, by = "county")%>% 
  mutate_at(vars("n_superfund"), ~replace(., is.na(.), 0))

combined_data_final=
  combined_data_final%>%
  mutate(
          site_score =  replace(site_score, is.na(site_score), 0),
          site_name = replace(site_name, is.na(site_name), "None"),
          superfund_site = if_else(n_superfund == '0',0, 1),
          cancer_population = round((cancer_crude_prev/100)*total_population),
          asthma_population = round((casthma_crude_prev/100)*total_population),
          smoking_population = round((csmoking_crude_prev/100)*total_population),
          depression_population = round((depression_crude_prev/100)*total_population),
          insurance_population = round((access2_crude_prev/100)*total_population),
          cancer_compared_national_average=if_else(cancer_crude_prev>'6.5', 1,0),
          )%>%
  distinct(county,.keep_all = TRUE)
```


##Outcome: Cancer Prevalence

The data provided is at the state-level, however it was converted to county-level for the purpose of this model. The outcome of interest is the number of individuals with cancer in a 1 year period by county differed by county.As this was count data, Poisson Regression with a log link function was utilized.No offset term was utilized in the data as count data is modeled.

#Crude Model: We evaluated 
```{r}
cancer_county_crude = 
  combined_data_final %>% 
  glm(cancer_population ~ superfund_site*county, data = ., family = poisson())%>%broom::tidy()%>%
  mutate(
    RR=exp(estimate),
    CI_lower=exp(estimate-1.96*std.error),
    CI_upper = exp(estimate+1.96*std.error)
  ) %>% 
  dplyr::select(term, RR, starts_with("CI")) 
  
fit_cancer_city_crude = 
  combined_data_final %>% 
  glm(cancer_population ~ superfund_site*city, data = ., family = poisson())%>%broom::tidy()%>%
  broom::tidy()%>%
    mutate(
    RR=exp(estimate),
    CI_lower=exp(estimate-1.96*std.error),
    CI_upper = exp(estimate+1.96*std.error)
  ) %>% 
  dplyr::select(term, RR, starts_with("CI")) %>% 

```

```{r}
fit_logistic_city = 
  combined_data_final %>%
   nest(data=-county)%>%
   mutate(data,models=map(data, ~glm(cancer_population ~ superfund_sites, data = .)),results = map(models, broom::tidy))%>%
   select(county,results)%>%
   unnest(cols = results)
```


#Confounder Selection: Analyze these values: 

`Depression`, 
```{r}
  combined_data_final %>% 
  glm(cancer_population ~depression_crude_prev, data = .,family = poisson(link='log'))%>%
  broom::tidy()

```

`Insurance_Status`, 
```{r}
  combined_data_final %>% 
  glm(cancer_population ~ insurance_population, data = .,family = poisson(link='log'))%>%
  broom::tidy()

```

`Smoking`
```{r}
  combined_data_final %>% 
  glm(cancer_population ~ csmoking_crude_prev, data = .,family = poisson(link='log'))%>%
  broom::tidy()

```

`site score`
```{r}
  combined_data_final %>% 
  glm(cancer_population ~ site_score, data = .,family = poisson(link='log'))%>%
  broom::tidy()

```

`Superfund Site`
```{r}
  combined_data_final %>% 
  glm(cancer_population ~ superfund_site, data = .,family = poisson(link='log'))%>%
  broom::tidy()

```

`Number of Superfund Sites`
```{r}
  combined_data_final %>% 
  glm(cancer_population ~ n_superfund, data = .,family = poisson(link='log'))%>%
  broom::tidy()

```

`total population`
```{r}
  combined_data_final %>% 
  glm(cancer_population ~ total_population, data = .,family = poisson(link='log'))%>%
  broom::tidy()

```

`city`
```{r}
  combined_data_final %>% 
  glm(cancer_population ~ city, data = .,family = poisson(link='log'))%>%
  broom::tidy()

```

#Adjusted Analysis
```{r}
fit_cancer_county_adjusted = 
  combined_data_final %>% 
  glm(cancer_population ~ insurance_population+depression_population+csmoking_crude_prev, data = .)%>%
  broom::tidy()

fit_cancer_city_adjusted = 
  combined_data_final %>% 
  glm(cancer_population ~ city*insurance_population*depression_population*csmoking_crude_prev, data = .,family = poisson(link='log'))%>%
  broom::tidy()
```

#Model Fit
```{r}
  combined_data_final%>% 
  glm(cancer_population ~ county, data = .,family = poisson(link='log'))%>% 
  dispersiontest(trafo = 1) %>% 
  broom::tidy() %>% 
  dplyr::select(estimate, p.value) 
```
