---
title: "Exploratory Analysis of Pennsylvania Health Outcomes"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(readr)
library(patchwork)
library(plotly)
library(lubridate)
library(stringr)

knitr::opts_chunk$set(echo = TRUE)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{=html}
<style type="text/css">

h1.title {
  text-align: center;
}

</style>
```
 

The main health outcomes of interest are asthma and cancer, since we hypothesize that residents who live in counties with Superfund sites are more likely to be exposed to hazardous waste which will increase their risk of cancer and asthma. Other variables we are interested in exploring are lack of health insurance, smoking, and depression along with the mean site score of the Superfunds in each county.

Primarily, we are interested in seeing if living in a county with high average Superfund site score would correlate with asthma or cancer prevalence.

 

```{r plot 4, message=FALSE, warning=FALSE}
#Data Cleaning to Extract 95% CIs 
health_outcomes = 
  read_csv("./data/PLACES_County_2022.csv") %>% 
  janitor::clean_names() %>% 
  rename(county = county_name, state = state_abbr) %>% 
  mutate(
    county = as.factor(county)
  ) %>% 
  select(state, county, casthma_crude_prev,casthma_crude95ci, cancer_crude_prev,cancer_crude95ci ) %>% 
  filter(state == "PA") %>%
  mutate(asthma_ci = str_remove_all(casthma_crude95ci, '[()]'),
         cancer_ci = str_remove_all(cancer_crude95ci, '[()]')) %>%
  select(-casthma_crude95ci,-cancer_crude95ci) %>%
  separate(asthma_ci, into = c("asthma_lower", "asthma_upper"), sep = ",") %>%
   separate(cancer_ci, into = c("cancer_lower", "cancer_upper"), sep = ",") %>%
  mutate(
    asthma_lower = as.numeric(asthma_lower),
    asthma_upper = as.numeric(asthma_upper),
    cancer_lower = as.numeric(cancer_lower),
    cancer_upper = as.numeric(cancer_upper)
  )
  
#Crude Asthma Prevalence Plot

asthma_plot =
ggplot(health_outcomes,
aes(x = reorder(county,-casthma_crude_prev), y = casthma_crude_prev, color = county)) + 
  geom_point() +
  geom_errorbar(aes(ymin = asthma_lower, ymax = asthma_upper)) +
  geom_hline(yintercept = 9.7, color = "red") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none") +
   labs(
    title = "Prevalence of Asthma Among Adults 18+ by County 2020",
    caption = "Red Line: national asthma prevalence calculated from national dataset",
    x = " ",
    y = "Crude Asthma Prevalence") 

ggplotly(asthma_plot)

```

\*\* Red Line: national asthma prevalence calculated from national dataset with all states.

The prevalence for asthma in the majority of Pennsylvania counties are above the calculated national average (9.7%). The county with the highest prevalence of asthma is Philadelphia County, while the county with the lowest is Montgomery County.  

```{r plot 5, message=FALSE, warning=FALSE}

#Crude Cancer Prevalence Plot

cancer_plot =
ggplot(health_outcomes,
aes(x = reorder(county,-cancer_crude_prev), y = cancer_crude_prev, color = county)) + 
  geom_point() +
  geom_errorbar(aes(ymin = cancer_lower, ymax = cancer_upper)) +
  geom_hline(yintercept = 5.6, color = "red") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none") +
  labs(
    title = "Prevalence of Cancer Among Adults 18+ by County 2020",
    caption = "Red Line: national cancer prevalence calculated from national dataset",
    x = " ",
    y = "Crude Cancer Prevalence") 

ggplotly(cancer_plot)
```

\*\* Red Line: national cancer prevalence calculated from CDC dataset with all states.

Again, from the plot above, the prevalence for cancer in the majority of Pennsylvania counties are above the calculated national average (5.6%). The county with the highest prevalence of cancer is Cameron County (9.5%), while the county with the lowest is Philadelphia County (5.5%).

 

```{r plot 6, message=FALSE, warning=FALSE}

#data cleaning 
health_prev = 
  read_csv("./data/PLACES_County_2022.csv") %>% 
  janitor::clean_names() %>% 
  rename(county = county_name, state = state_abbr) %>% 
  mutate(
    county = as.factor(county)
  )  %>% 
  filter(state == "PA") %>% 
  select(state, county,casthma_crude_prev,cancer_crude_prev,
         csmoking_crude_prev,depression_crude_prev,access2_crude_prev)

superfund_sites =
  read_csv("./data/superfund_NPL_sites.csv") %>% 
  janitor::clean_names() %>% 
  select(-region_id,-construction_completion_number, -site_listing_narrative, -site_progress_profile, -ends_with("notice"), -restoration_fr_notice_jumper_page, -noid_date, -deletion_date) %>% 
  filter(status == "NPL Site")

superfund_score =
  superfund_sites %>% 
  group_by(county) %>% 
  summarize(
    n_superfund = n(),
    mean_score = mean(site_score)) %>% 
  mutate(
    county = as.factor(county)) %>% 
  arrange(desc(mean_score)) 

health_score = 
  full_join(superfund_score, health_prev, by = "county") %>% 
  drop_na()

#Asthma vs Smoking
  
Asthma_Smoking =
health_prev %>% 
  select(county,csmoking_crude_prev,casthma_crude_prev) %>%
  ggplot(
    aes(x = csmoking_crude_prev,y = casthma_crude_prev)) + 
  geom_point(aes(color = county)) +
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Crude Smoking Prevalence ",
    y = "Crude Asthma Prevalence"
  ) 

#Asthma vs Depression

Asthma_Depression =
health_prev %>% 
  select(county,depression_crude_prev,casthma_crude_prev) %>%
  ggplot(
    aes(x = depression_crude_prev,y = casthma_crude_prev)) + 
  geom_point(aes(color = county)) +
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Crude Depression Prevalence ",
    y = "Crude Asthma Prevalence"
  ) 

#Asthma vs Lack of Insurance Access 

Asthma_Access =
health_prev %>% 
  select(county,access2_crude_prev,casthma_crude_prev) %>%
  ggplot(
    aes(x = access2_crude_prev,y = casthma_crude_prev)) + 
  geom_point(aes(color = county)) +
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Crude Prevalence of Lack of Insurance",
    y = "Crude Asthma Prevalence"
  ) 



#Asthma vs Mean Site Score

Asthma_Score =
  health_score %>% 
   select(county,mean_score,casthma_crude_prev) %>%
  ggplot(
    aes(x = mean_score,y = casthma_crude_prev)) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_point(aes(color = county)) +
   theme(legend.position = "none") + 
  labs(
    x = "Mean of Superfund Site Score In Each County",
    y = "Crude Asthma Prevalence"
  ) 

asthma_predict_panel = (Asthma_Smoking + Asthma_Depression) / (Asthma_Access + Asthma_Score)

asthma_predict_panel + 
  plot_annotation(
    title = "Association Between Asthma Prevalence and Given Predictors",
    subtitle = " County Level Data") + 
  theme(
    legend.position = "bottom"
  ) + 
  plot_layout(guides = "collect")
```

 

From these four scatter plots with best fit lines grouped by asthma as the dependent variable, smoking prevalence, depression prevalence, and lack of insurance, all have a positive relationship with asthma prevalence on the county level. Mean Superfund site score seems to be more randomly distributed, with a slight downward trend that is likely influenced by the one outlier.

```{r plot 7, message=FALSE, warning=FALSE}
Cancer_Smoking =
health_prev %>% 
  select(county,csmoking_crude_prev,cancer_crude_prev) %>%
  ggplot(
    aes(x = csmoking_crude_prev,y = cancer_crude_prev)) + 
  geom_point(aes(color = county)) + 
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Crude Smoking Prevalence ",
    y = "Crude Cancer Prevalence"
  ) 


Cancer_Depression =
health_prev %>% 
  select(county,depression_crude_prev,cancer_crude_prev) %>%
  ggplot(
    aes(x = depression_crude_prev,y = cancer_crude_prev)) + 
  geom_point(aes(color = county)) +
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Crude Depression Prevalence ",
    y = "Crude Cancer Prevalence"
  ) 

Cancer_Access =
health_prev %>% 
  select(county,access2_crude_prev,cancer_crude_prev) %>%
  ggplot(
    aes(x = access2_crude_prev,y = cancer_crude_prev)) + 
  geom_point(aes(color = county)) +
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Crude Prevalence of Lack of Insurance",
    y = "Crude Cancer Prevalence"
  ) 

Cancer_Score =
  health_score %>% 
   select(county,mean_score,cancer_crude_prev) %>%
  ggplot(
    aes(x = mean_score,y = cancer_crude_prev)) + 
  geom_point(aes(color = county)) + 
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Mean of Superfund Site Score In Each County",
    y = "Crude Cancer Prevalence"
  ) 

cancer_predict_panel = (Cancer_Smoking + Cancer_Depression) / (Cancer_Access + Cancer_Score)

cancer_predict_panel + 
  plot_annotation(
    title = "Association Between Cancer Prevalence and Given Predictors",
    subtitle = " County Level Data") + 
  theme(
    legend.position = "bottom"
  ) + 
  plot_layout(guides = "collect")
```

For Cancer Prevalence as the outcome, smoking and depression have a slight positive correlation with ashtma prevalence on the county level. On the other hand, lack of insurance and mean superfund site score seem to have a more random distribution with cancer. The best fit line for these two outcomes show a slight negative correlation but it is likely influenced by the outliers.

```{r plot 8, message=FALSE, warning=FALSE}
Access_Score =
  health_score %>% 
   select(county,mean_score,access2_crude_prev) %>%
  ggplot(
    aes(x = access2_crude_prev,y = mean_score)) + 
  geom_point(aes(color = county)) + 
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Prevalence of Lack of Health Insurance",
    y = "Mean County Superfund Site Score"
  ) +
  ggtitle("Association Between Mean Site Score and Lack of Insurance Prevalence")
Access_Score
```

 

There appears to be very low correlation between lack of health insurance and mean county Superfund site core at the county level.

```{r plot 9, message=FALSE, warning=FALSE}
no_superfund = 
  full_join(superfund_score, health_prev, by = "county") %>%
  filter(is.na(n_superfund)) %>%
  select(n_superfund,casthma_crude_prev,cancer_crude_prev,access2_crude_prev) %>%
  mutate(has_superfund = "No") %>%
  select(-n_superfund) %>%
  mutate(has_superfund = as.factor(has_superfund))

yes_superfund =
  health_score %>%
  select(n_superfund,casthma_crude_prev,cancer_crude_prev,access2_crude_prev) %>%
    mutate(has_superfund = "Yes") %>%
    select(-n_superfund) %>%
    mutate(has_superfund = as.factor(has_superfund))

health_superfund =
  rbind(no_superfund,yes_superfund) %>%
  pivot_longer(
    casthma_crude_prev:access2_crude_prev,
    names_to = "outcome_type", 
    values_to = "crude_prevalence")
  

asthma_superfund =
   health_superfund %>%
   plot_ly(y = ~crude_prevalence, x = ~has_superfund ,
           type = "box", color = ~outcome_type) %>% 
  layout(boxmode = "group") %>% 
  layout(
    xaxis = list(title = "Presence of Superfund in County"),
   title = "Comparison of County Level Cancer, Asthma Prevalence, and Lack of Health Insurance By Superfund Presence",
   yaxis = list(title = "Crude Prevalence")) 

asthma_superfund 
```

Counties that have Superfund sites located in their counties compared to counties that did not have any Superfund sites in their counties appeared to have similar median prevalence values across all three variables of asthma, cancer, and lack of health insurance prevalence. Counties with Superfund sites do have a larger range for their prevalence of lack of health insurance.
