---
title: "Exploratory Analysis of Pennsylvania Health Outcomes"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(readr)
library(patchwork)
library(plotly)
library(lubridate)
library(stringr)

knitr::opts_chunk$set(echo = TRUE)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

<style type="text/css">

h1.title {
  text-align: center;
}

</style>

&nbsp;

The main health outcomes of interest are asthma and cancer, since we hypothesize that residents who live in counties with Superfund sites are more likely to be exposed to hazardous waste which will increase their risk of cancer and asthma. Other variables we are interested in exploring are lack of health insurance, smoking, and depression along with the mean site score of the Superfunds in each county.

&nbsp;
```{r plot 4, message=FALSE, warning=FALSE}
#Data Cleaning to Extract 95% CIs 
health_outcomes = 
  read_csv("./data/PLACES_County_2022.csv") %>% 
  janitor::clean_names() %>% 
  rename(county = county_name, state = state_abbr) %>% 
  mutate(
    county = as.factor(county)
  ) %>% 
  select(state, county, casthma_crude_prev,casthma_crude95ci, cancer_crude_prev,cancer_crude95ci ) %>% 
  filter(state == "PA") %>%
  mutate(asthma_ci = str_remove_all(casthma_crude95ci, '[()]'),
         cancer_ci = str_remove_all(cancer_crude95ci, '[()]')) %>%
  select(-casthma_crude95ci,-cancer_crude95ci) %>%
  separate(asthma_ci, into = c("asthma_lower", "asthma_upper"), sep = ",") %>%
   separate(cancer_ci, into = c("cancer_lower", "cancer_upper"), sep = ",") %>%
  mutate(
    asthma_lower = as.numeric(asthma_lower),
    asthma_upper = as.numeric(asthma_upper),
    cancer_lower = as.numeric(cancer_lower),
    cancer_upper = as.numeric(cancer_upper)
  )
  
#Crude Asthma Prevalence Plot

asthma_plot =
ggplot(health_outcomes,
aes(x = reorder(county,-casthma_crude_prev), y = casthma_crude_prev, color = county)) + 
  geom_point() +
  geom_errorbar(aes(ymin = asthma_lower, ymax = asthma_upper)) +
  geom_hline(yintercept = 9.7, color = "red") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none") +
   labs(
    title = "Prevalence of Asthma Among Adults 18+ by County 2020",
    caption = "Red Line: national asthma prevalence calculated from national dataset",
    x = " ",
    y = "Crude Asthma Prevalence") 

ggplotly(asthma_plot)

```
** Red Line: national asthma prevalence calculated from national dataset with all states.

```{r plot 5, message=FALSE, warning=FALSE}

#Crude Cancer Prevalence Plot

cancer_plot =
ggplot(health_outcomes,
aes(x = reorder(county,-cancer_crude_prev), y = cancer_crude_prev, color = county)) + 
  geom_point() +
  geom_errorbar(aes(ymin = cancer_lower, ymax = cancer_upper)) +
  geom_hline(yintercept = 5.6, color = "red") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "none") +
  labs(
    title = "Prevalence of Cancer Among Adults 18+ by County 2020",
    caption = "Red Line: national cancer prevalence calculated from national dataset",
    x = " ",
    y = "Crude Cancer Prevalence") 

ggplotly(cancer_plot)
```
** Red Line: national cancer prevalence calculated from CDC dataset with all states.

```{r plot 6, message=FALSE, warning=FALSE}

#data cleaning 
health_prev = 
  read_csv("./data/PLACES_County_2022.csv") %>% 
  janitor::clean_names() %>% 
  rename(county = county_name, state = state_abbr) %>% 
  mutate(
    county = as.factor(county)
  )  %>% 
  filter(state == "PA") %>% 
  select(state, county,casthma_crude_prev,cancer_crude_prev,
         csmoking_crude_prev,depression_crude_prev,access2_crude_prev)

superfund_sites =
  read_csv("./data/superfund_NPL_sites.csv") %>% 
  janitor::clean_names() %>% 
  select(-region_id,-construction_completion_number, -site_listing_narrative, -site_progress_profile, -ends_with("notice"), -restoration_fr_notice_jumper_page, -noid_date, -deletion_date) %>% 
  filter(status == "NPL Site")

superfund_score =
  superfund_sites %>% 
  group_by(county) %>% 
  summarize(
    n_superfund = n(),
    mean_score = mean(site_score)) %>% 
  mutate(
    county = as.factor(county)) %>% 
  arrange(desc(mean_score)) 

health_score = 
  full_join(superfund_score, health_prev, by = "county") %>% 
  drop_na()

#Asthma vs Smoking
  
Asthma_Smoking =
health_prev %>% 
  select(county,csmoking_crude_prev,casthma_crude_prev) %>%
  ggplot(
    aes(x = csmoking_crude_prev,y = casthma_crude_prev)) + 
  geom_point(aes(color = county)) +
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Crude Smoking Prevalence ",
    y = "Crude Asthma Prevalence"
  ) 

#Asthma vs Depression

Asthma_Depression =
health_prev %>% 
  select(county,depression_crude_prev,casthma_crude_prev) %>%
  ggplot(
    aes(x = depression_crude_prev,y = casthma_crude_prev)) + 
  geom_point(aes(color = county)) +
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Crude Depression Prevalence ",
    y = "Crude Asthma Prevalence"
  ) 

#Asthma vs Lack of Insurance Access 

Asthma_Access =
health_prev %>% 
  select(county,access2_crude_prev,casthma_crude_prev) %>%
  ggplot(
    aes(x = access2_crude_prev,y = casthma_crude_prev)) + 
  geom_point(aes(color = county)) +
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Crude Prevalence of Lack of Insurance",
    y = "Crude Asthma Prevalence"
  ) 



#Asthma vs Mean Site Score

Asthma_Score =
  health_score %>% 
   select(county,mean_score,casthma_crude_prev) %>%
  ggplot(
    aes(x = mean_score,y = casthma_crude_prev)) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_point(aes(color = county)) +
   theme(legend.position = "none") + 
  labs(
    x = "Mean of SuperFund Site Score In Each County",
    y = "Crude Asthma Prevalence"
  ) 

asthma_predict_panel = (Asthma_Smoking + Asthma_Depression) / (Asthma_Access + Asthma_Score)

asthma_predict_panel + 
  plot_annotation(
    title = "Association Between Asthma Prevalence and Given Predictors",
    subtitle = " County Level Data") + 
  theme(
    legend.position = "bottom"
  ) + 
  plot_layout(guides = "collect")
```

```{r plot 7, message=FALSE, warning=FALSE}
Cancer_Smoking =
health_prev %>% 
  select(county,csmoking_crude_prev,cancer_crude_prev) %>%
  ggplot(
    aes(x = csmoking_crude_prev,y = cancer_crude_prev)) + 
  geom_point(aes(color = county)) + 
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Crude Smoking Prevalence ",
    y = "Crude Cancer Prevalence"
  ) 


Cancer_Depression =
health_prev %>% 
  select(county,depression_crude_prev,cancer_crude_prev) %>%
  ggplot(
    aes(x = depression_crude_prev,y = cancer_crude_prev)) + 
  geom_point(aes(color = county)) +
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Crude Depression Prevalence ",
    y = "Crude Cancer Prevalence"
  ) 

Cancer_Access =
health_prev %>% 
  select(county,access2_crude_prev,cancer_crude_prev) %>%
  ggplot(
    aes(x = access2_crude_prev,y = cancer_crude_prev)) + 
  geom_point(aes(color = county)) +
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Crude Prevalence of Lack of Insurance",
    y = "Crude Cancer Prevalence"
  ) 

Cancer_Score =
  health_score %>% 
   select(county,mean_score,cancer_crude_prev) %>%
  ggplot(
    aes(x = mean_score,y = cancer_crude_prev)) + 
  geom_point(aes(color = county)) + 
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Mean of SuperFund Site Score In Each County",
    y = "Crude Cancer Prevalence"
  ) 

cancer_predict_panel = (Cancer_Smoking + Cancer_Depression) / (Cancer_Access + Cancer_Score)

cancer_predict_panel + 
  plot_annotation(
    title = "Association Between Xancer Prevalence and Given Predictors",
    subtitle = " County Level Data") + 
  theme(
    legend.position = "bottom"
  ) + 
  plot_layout(guides = "collect")
```
```{r plot 8, message=FALSE, warning=FALSE}
Access_Score =
  health_score %>% 
   select(county,mean_score,access2_crude_prev) %>%
  ggplot(
    aes(x = access2_crude_prev,y = mean_score)) + 
  geom_point(aes(color = county)) + 
  geom_smooth(method = "lm", se = FALSE) +
   theme(legend.position = "none") + 
  labs(
    x = "Prevalence of Lack of Health Insurance",
    y = "Mean County Superfund Site Score"
  ) +
  ggtitle("Association Between Mean Site Score and Lack of Insurance Prevalence")
Access_Score
```

```{r plot 9, message=FALSE, warning=FALSE}
no_superfund = 
  full_join(superfund_score, health_prev, by = "county") %>%
  filter(is.na(n_superfund)) %>%
  select(n_superfund,casthma_crude_prev,cancer_crude_prev,access2_crude_prev) %>%
  mutate(has_superfund = "No") %>%
  select(-n_superfund) %>%
  mutate(has_superfund = as.factor(has_superfund))

yes_superfund =
  health_score %>%
  select(n_superfund,casthma_crude_prev,cancer_crude_prev,access2_crude_prev) %>%
    mutate(has_superfund = "Yes") %>%
    select(-n_superfund) %>%
    mutate(has_superfund = as.factor(has_superfund))

health_superfund =
  rbind(no_superfund,yes_superfund) %>%
  pivot_longer(
    casthma_crude_prev:access2_crude_prev,
    names_to = "outcome_type", 
    values_to = "crude_prevalence")
  

asthma_superfund =
   health_superfund %>%
   plot_ly(y = ~crude_prevalence, x = ~has_superfund ,
           type = "box", color = ~outcome_type) %>% 
  layout(boxmode = "group") %>% 
  layout(
    xaxis = list(title = "Presence of Superfund in County"),
   title = "Comparison of County Level Cancer, Asthma Prevalence, and Lack of Health Insurance By Superfund Presence",
   yaxis = list(title = "Crude Prevalence")) 

asthma_superfund 
 
```